---
title: "03_preprocessing"
author: "Daphne Garcia"
date: "2025-03-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

Goals of this file
The goal is to make a phylogenetic tree!

Load in preprocessed phyloseq object
Create ASV fasta file from the phyloseq object
Align the 16S sequences from fasta file with MAFT
Create a tree with FastTree2

##Set seed
```{r set seed}

set.seed(238428)
```

## Load packages
```{r load environment}

#install phytools and ggtree
remotes::install_github("liamrevell/phytools")
devtools::install_github("GuangchuangYu/ggtree")


pacman::p_load(phytools, phyloseq, ggtree, tidyverse, install = FALSE)


# load physeq
load("data/03_PreProcessing/raw_preprocessed_physeq.RData")
load("data/03_PreProcessing/raw_preprocessed_physeq.RData")
getwd()
```

#Write ASV Fasta File
```{r create_preprocessed_fasta}

# Pull out ASV sequences and ASV names
asv_seq_df <- 
  raw_preprocessed_physeq@tax_table %>% 
  data.frame() %>% 
  dplyr::select(ASV, ASVseqs)

#View(asv_seq_df)

#Add the ">" to make fasta header
asv_seq_df$ASV <- paste0(">", asv_seq_df$ASV)

#Create fasta object
asv_seq_fasta <- c(rbind(asv_seq_df$ASV, asv_seq_df$ASVseqs))

#write to a fasta file
write(asv_seq_fasta,
      file = "data/04_Phylogenetic_Tree/preprocessed_ASVs.fasta")

```

#Align the 16S sequences from fasta file with MAFT
- `engine.opts = -l: this gives us original terminal base environment`
```{r run-maft, engine='bash', engine.opts = '-l}
#write bash code to run MAFT alignment

#first, provide path to MAFT on server (found on biohpc website, MAFFT)
---
title: "03_preprocessing"
author: "Daphne Garcia"
date: "2025-03-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

Goals of this file
The goal is to make a phylogenetic tree!

Load in preprocessed phyloseq object
Create ASV fasta file from the phyloseq object
Align the 16S sequences from fasta file with MAFT
Create a tree with FastTree2

##Set seed
```{r set seed}

set.seed(238428)
```

## Load packages
```{r load environment}

#install phytools and ggtree
remotes::install_github("liamrevell/phytools")
devtools::install_github("GuangchuangYu/ggtree")


pacman::p_load(phytools, phyloseq, ggtree, tidyverse, install = FALSE)


# load physeq
load("data/03_PreProcessing/raw_preprocessed_physeq.RData")
load("data/03_PreProcessing/raw_preprocessed_physeq.RData")
getwd()
```

#Write ASV Fasta File
```{r create_preprocessed_fasta}

# Pull out ASV sequences and ASV names
asv_seq_df <- 
  raw_preprocessed_physeq@tax_table %>% 
  data.frame() %>% 
  dplyr::select(ASV, ASVseqs)

#View(asv_seq_df)

#Add the ">" to make fasta header
asv_seq_df$ASV <- paste0(">", asv_seq_df$ASV)

#Create fasta object
asv_seq_fasta <- c(rbind(asv_seq_df$ASV, asv_seq_df$ASVseqs))

#write to a fasta file
write(asv_seq_fasta,
      file = "data/04_Phylogenetic_Tree/preprocessed_ASVs.fasta")

```

#Align the 16S sequences from fasta file with MAFT
- `engine.opts = -l: this gives us original terminal base environment`
- have to run with ctrl + alt + enter
```{r run-maft, engine='bash', engine.opts='-l'}
#write bash code to run MAFT alignment

#first, provide path to MAFT on server (found on biohpc website, MAFFT)
export PATH=/programs/mafft/bin:$PATH

ls
# change directories to provide fasta file we made above
cd data/04_Phylogenetic_Tree

pwd

#set a seed for consistency and reproducibility
RANDOM=238428

#MAFFT already knows that we have a nucleotide alignment
/programs/mafft/bin/mafft --auto preprocessed_ASVs.fasta > MAFFT_aligned_ASVs.fasta

# Change back to the project directory 
cd ../../
pwd

```

#FastTree2
```{r run_fasttree, engine='bash', engine.opts='-l'}

# provide export path to FastTree2
export PATH=/programs/FastTree-2.1.11/FastTree:$PATH ##This didn't work
export PATH=/programs/FastTree-2.1.11:$PATH   ## I got this from biohpc, worked

#change directory into alignment file folder(this is same as html guide, but not what we did in class)
cd data/04_Phylogenetic_Tree/

#run FastTree to generate phylogenetic tree
#parameters: 
  # -nt: indicates its a nuccleotide alignment
  # -gtr = generalized time reversible substitution model
  # -fastest = speed up the model, reduce memory usage(recommended dataset that have over 50k)
  # -log = output a log file
  # input our alignment file
  # specify the output tree file
  
FastTree -nt -gtr -fastest -log FastTree.log MAFFT_aligned_ASVs.fasta > ASVs_unrooted.tree

#change back to  main directory
cd ../../
pwd
```

# 
